### 2020-02-05 README.Rmd
[docs] Rename project-wide
# |<----  Using a Maximum Of 50 Characters  ---->|


# Explain why this change is being made
# |<----   Try To Limit Each Line to a Maximum Of 72 Characters   ---->|

# Provide links or keys to any relevant tickets, articles or other resources
# Example: Github issue #23

# --- COMMIT END ---
# Type can be 
#    feat (new feature)
#    fix (bug fix)
#    docs (changes to documentation)
#    style (formatting, missing semi colons, etc; no code change)
#    refactor (refactoring production code)
#    test (adding missing tests, refactoring tests; no production code cha

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

# Please put your title here to include it in the file below.
Title <- "An R-based framework for conducting parametric analysis on building energy models"
```

# eplusr + epluspar

<!-- badges: start -->
[![R build status](https://github.com/ideas-lab-nus/epluspar-paper/workflows/build-paper/badge.svg)](https://github.com/ideas-lab-nus/epluspar-paper/actions)
<!-- badges: end -->

This repository contains the data and code for our paper:

> Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r
> Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at
> <https://doi.org/xxx/xxx>

### How to download or install

You can download the compendium as a zip from from this URL:
<https://github.com/ideas-lab-nus/epluspar-paper/archive/master.zip>

Or you can install this compendium as an R package, epluspar-paper, from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("remotes")
remotes::install_github("ideas-lab-nus/epluspar-paper")
```

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution
requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our
[contributor guidelines](.github/CONTRIBUTING.md). Please note that this project
is released with a [Contributor Code of Conduct](.github/CONDUCT.md). By participating
in this project you agree to abide by its terms.